---
title: "fssg_intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fssg_intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

fssg is a small package focused on allowing users to quickly test various parametric survival models in one line of code. This package was originally inspired some of my other research on Broad Genomic Profiling, where tens to hundreds of genomic mutations are tested in one single batch. This approach can be referred to in a few ways, but may be colloquially referred to as a "shotgun" approach. Hence, fssg ~ FlexSurv ShotGun.

To get started with fssg, we'll simply load our package, which will also grab the main dependency flexsurv.

```{r setup}
library(fssg)
library(flexsurv)
```

The main powerhouse behind this package is the parametric survival modeling package "flexsurv". flexsurv provides a straightforward way to build parametric survival distributions with many of the most commonly used distributions (such as weibull, gamma, etc.). This is a very useful tool, but is limited in parametric distribution options. Thankfully, the authors of flexsurv allow users to declare and use their own custom defined distribution as well. This means that as long as we know what we're doing, we can use whichever distribution we'd like!


## Custom Distributions

Included in this package are some functions to streamline the creation of custom distributions for use inside flexsurv's regression functions. "fssg_dist" is a function which builds the distribution object when supplied with the necessary parameters. 

What does our custom distribution need?

1. A name! Used in the fssg backend, but technically isn't that important if you're not doing batch testing via fssg.
2. A vector of parameters. This lets flexsurv know what parameters need to be initialized and optimized in the fitting process.
3. An indicator for the main variable we want to change depending on the covariates supplied. flexsurv calls this the 'location' variable, as it originally focused mostly on location-scale distributions, but we can specify any of the distribution parameters here. Please note, that this can only be one parameter in flexsurv, but we can allow other parameters to vary by supplying a list into the "anc" parameter in flexsurvreg. See ?flexsurvreg for more information on this.
4. Transformation functions that change our parameter domains into the real number line for better optimization. flexsurv also request for the inverse functions to allow for back transformation to the parameter scale. These are supplied to the distribution object as a list of functions. Most often, when parameters must be greater than 0, the transforms function is something like "log", while the inverse transformation is "exp".
5. Initial values. The optimization process requires a starting point. This is often the trickiest part of specifying a distribution, as poor initial values will cause the optimization to fail. To avoid these problems, it's recommend to have your inits supplied as function which is dependent on the survival times. This allows the initial values to be dynamically determined by the actual shape of your data, and lowers the chance of bad initial values. 
6. Distribution functions. flexsurv actually allows to be specified in one of two primary ways: with a density and distribution function ('d', 'p'), or with a Hazard and Cumulative Hazard function ('h','H'). It also helps to include the q function when possible. This package also includes some helper functions to construct these distribution functions based off a 'd' and 'p' function (See survivify).

This package contains some custom distribution functions for the gamma-gompertz distribution (e.g. dgamgomp, pgamgomp, etc.). These distributions functions are specified in a similar fashion to the native R distribution functions like pnorm, dnorm, qnorm etc.

```{r}
fssg_gamgomp <- list(
    name='gamgomp',                       # a simple name
    pars=c('b','sigma','beta'),           # scale, shape, shape
    location='b',                         # name of the parameter we want to vary 
    transforms=c(log,log,log),            # transformation functions for each of our parameters
    inv.transforms=c(exp,exp,exp),        # back-transformation functions for each of our parameters
    inits=function(t){c(1/max(t), 1, 1)}, # function that constructs the initial parameter estimates for optimization
    d = dgamgomp,                         # density function! fssg comes with this density function for Gamma-Gompertz by default
    p = pgamgomp,                         # accompanying distribution function.
    q = quantilify(pgamgomp),             # quantile function! used in some of the other fssg package functions
    h = hazardify(dgamgomp, pgamgomp),    # optional hazard function
    H = cumhazardify(pgamgomp),           # optional cumulative hazard function
    fullname='gamma_gompertz'             # secondary name which is typically used in the back-end of fssg to better label outputs
  )
```

We can then use this distribution object directly with the normal flexsurvreg function. It's recommended that you specify the "dfns" parameter explicitly, as it will attempt to use globally declared d and p variables based on your distribution name otherwise. 

```{r}
data('pseudo',package='fssg')  # sample data set provided in the package

flexsurvreg(
  formula = Surv(time, death) ~ 1, 
  data = pseudo, 
  dist = fssg_gamgomp, 
  dfns = list(fssg_gamgomp$d, fssg_gamgomp$p)
) -> f1

print(f1)
plot(f1)
```

To save users some time, we've already specified a large number of custom distributions, which can be cleanly accessed using fssg_dist_list(). This function creates a large list of each included distribution object. If you want to grab a specific distribution object, you can directly grab it using get_fssg_dist().


```{r}
get_fssg_dist('gamma_gompertz')
```

We can also use the fssg function to build an individual flexsurvreg model by giving it individual model names! More on this function later.

```{r}
fssg(
  formula = Surv(time, death) ~ 1, 
  data = pseudo, 
  model='gamma_gompertz'
) -> model1

model1$models$gamma_gompertz
```


## The `fssg` Function

The titular `fssg` function provides a one-line way to test the fit of multiple distributions on your data. Simply hand it your model formula and data, and watch it run. By default, the function will attempt to build all flexsurv and fssg parametric models on your data. Progress will be printed by default, and any models that failed will list the corresponding error. The function always prints a summary of all models run, and by default returns a list with two items: the summary table as a data.frame, and a list containing each successful model.

```{r}
fssg(survival::Surv(time, status) ~ x, data = survival::aml) -> output1
```

In many cases, the user may only want to try a subset of the available models. This can be done by giving `fssg` only the names of the distributions that you want. These should correspond to either the `name` or `fullname` fields found in the package distribution objects.

```{r}
fssg(
  survival::Surv(time, death) ~ gender, 
  data = pseudo, 
  models = c('gamma_gompertz','gompertz','gumbel','gumbel_T2')
) 
```


## Failed models

Due to the large number of distributions considered in the `fssg` function, it is likely that one or more of them will fail to work with the provided data. This is typically due to one of three reasons:

1. Poor initial values
  * Error's such as "initial value in 'vmmin' is not finite" are evidence of this
2. Optimized into invalid values
  * Error's such as "non-finite finite-difference value" are evidence of successful initial values, but failure's in the optimization process. This happens when `optim` perturbs a parameters, but the new perturbed value leads to invalid results such as `NaN` or `Inf`.
3. Failure to identify parameters
  * Error's like "system is exactly singular" are indicative of hessian computation issues. Can occur when the parameters are not estimable for the given data. 
  

The first 

### Initial Values

fssg includes a basic diagnostic function for checking whether the `init` function for a distribution will fail or not based on some data. This checks if the initial distribution parameters provided by `init` will allow the distribution function to return valid values for each observed time values. 

```{r}
check_inits(times = pseudo$time,  get_fssg_dist('gumbel'))
check_inits(times = pseudo$time,  get_fssg_dist('gumbel_T2'))
check_inits(times = pseudo$time,  get_fssg_dist('lomax'))
```

## Individual Models

Each model created by fssg can be interacted with in the same way you would interact with flexsurv models. Here's a quick example of how to look into a specific model using the gumbel distribution.

```{r}
fssg(survival::Surv(time, status) ~ age + sex, 
  data = survival::cancer, 
  models = c('gamma_gompertz','gompertz','gumbel','gumbel_T2'), progress = F) -> output

output$models$gumbel
```
Sex appears to be a significant variable here, as the confidence interval does not contain zero. Let's see what this exp(est) of 1.56 amounts to visually.

```{r}
plot(
  output$models$gumbel, 
  newdata = data.frame(sex=c(1,2), age=70), col=c('blue','red'),
  type='survival',
  lwd=2,
)
legend(
  'topright',
  legend = c('Male','Female'),
  col =c('blue','red'),
  lwd=2
)
```

It looks like males have a considerably higher hazard rate over time according to this model. 

On the other hand, age is not significantly associated with survival time. 

```{r}
plot(
  output$models$gumbel, 
  newdata = data.frame(sex=c(1), age=c(65,80)), col=c('blue','red'),
  type='survival',
  lwd=2
)
legend(
  'topright',
  legend = c('Age 65','Age 80'),
  col =c('blue','red'),
  lwd=2
)
```

This makes sense, as the two survival lines appear to be very similar. 


## Fit Statistics 

By default, fssg provides AIC, BIC, and Log Likelihood for the models in the summary table provided.

```{r}
model1$summary
```

Included in this package is a function that compiles a wide variety of fit statistics for a model, including concordance, AUCs, Mean Absolute Error (MAE), Integrated Absolute Error (IAE), Integrated Square Error (ISE), Brier Scores, and optionally the very slow to calculate Integrated Brier Scores.

```{r}
get_fit_stats(model1$models$gamma_gompertz) -> fitstats

fitstats
```

More specifics on these metrics can be found in the vignette "Fit_Statistics".



