---
title: "fssg_intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fssg_intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

fssg is a small package focused on allowing users to quickly test various parametric survival models in one line of code. This package was originally inspired some of my other research on Broad Genomic Profiling, where tens to hundreds of genomic mutations are tested in one single batch. This approach can be referred to in a few ways, but may be colloquially referred to as a "shotgun" approach. Hence, fssg ~ FlexSurv ShotGun.

To get started with fssg, we'll simply load our package, which will also grab the main dependency flexsurv.

```{r setup}
library(fssg)
library(flexsurv)
library(survival)
```

The main powerhouse behind this package is the parametric survival modeling package "flexsurv". flexsurv provides a straightforward way to build parametric survival distributions with many of the most commonly used distributions (such as weibull, gamma, etc.). This is a very useful tool, but is limited in parametric distribution options. Thankfully, the authors of flexsurv allow users to declare and use their own custom defined distribution as well. This means that as long as we know what we're doing, we can use whichever distribution we'd like!


## Custom Distributions

Included in this package are some functions to streamline the creation of custom distributions for use inside flexsurv's regression functions. "fssg_dist" is a function which builds the distribution object when supplied with the necessary parameters. 

What does our custom distribution need?

1. A name! Used in the fssg backend, but technically isn't that important if you're not doing batch testing via fssg.
2. A vector of parameters. This lets flexsurv know what parameters need to be initialized and optimized in the fitting process.
3. An indicator for the main variable we want to change depending on the covariates supplied. flexsurv calls this the 'location' variable, as it originally focused mostly on location-scale distributions, but we can specify any of the distribution parameters here. Please note, that this can only be one parameter in flexsurv, but we can allow other parameters to vary by supplying a list into the "anc" parameter in flexsurvreg. See ?flexsurvreg for more information on this.
4. Transformation functions that change our parameter domains into the real number line for better optimization. flexsurv also request for the inverse functions to allow for back transformation to the parameter scale. These are supplied to the distribution object as a list of functions. Most often, when parameters must be greater than 0, the transforms function is something like "log", while the inverse transformation is "exp".
5. Initial values. The optimization process requires a starting point. This is often the trickiest part of specifying a distribution, as poor initial values will cause the optimization to fail. To avoid these problems, it's recommend to have your inits supplied as function which is dependent on the survival times. This allows the initial values to be dynamically determined by the actual shape of your data, and lowers the chance of bad initial values. 
6. Distribution functions. flexsurv actually allows to be specified in one of two primary ways: with a density and distribution function ('d', 'p'), or with a Hazard and Cumulative Hazard function ('h','H'). It also helps to include the q function when possible. This package also includes some helper functions to construct these distribution functions based off a 'd' and 'p' function (See survivify).

Let's see an example!



This package contains some custom distribution functions for the gamma-gompertz distribution (e.g. dgamgomp, pgamgomp, etc.). These distributions functions are specified in a similar fashion to the native R distribution functions like pnorm, dnorm, qnorm etc.

```{r}
fssg_gamgomp <- list(
    name='gamgomp',                       # a simple name
    pars=c('b','sigma','beta'),           # scale, shape, shape
    location='b',                         # name of the parameter we want to vary 
    transforms=c(log,log,log),            # transformation functions for each of our parameters
    inv.transforms=c(exp,exp,exp),        # back-transformation functions for each of our parameters
    inits=function(t){c(1/max(t), 1, 1)}, # function that constructs the initial parameter estimates for optimization
    d = dgamgomp,                         # density function! fssg comes with this density function for Gamma-Gompertz by default
    p = pgamgomp,                         # accompanying distribution function.
    q = quantilify(pgamgomp),             # quantile function! used in some of the other fssg package functions
    h = hazardify(dgamgomp, pgamgomp),    # optional hazard function
    H = cumhazardify(pgamgomp),           # optional cumulative hazard function
    fullname='gamma_gompertz'             # secondary name which is typically used in the back-end of fssg to better label outputs
  )
```

We can then use this distribution object directly with the normal flexsurvreg function. It's recommended that you specify the "dfns" parameter explicitly, as it will attempt to use globally declared d and p variables based on your distribution name otherwise. 


```{r}
flexsurvreg(
  formula = Surv(time, status) ~ 1, 
  data=survival::aml, 
  dist = fssg_gamgomp, 
  dfns = list(fssg_gamgomp$d, fssg_gamgomp$p)
) -> f1

print(f1)
plot(f1)
```

To save users some time, we've already specified a large number of custom distributions, which can be cleanly accessed using fssg_dist_list(). This function creates a large list of each included distribution object. If you want to grab a specific distribution object, you can directly grab it using get_fssg_dist().


```{r}
get_fssg_dist('gamma_gompertz')
```

We can also use the fssg function to build an individual flexsurvreg model by giving it individual model names! More on this function later.

```{r}
fssg(
  formula = Surv(time, status) ~ 1, 
  data=survival::aml, 
  model='gamma_gompertz'
)$models$gamma_gompertz
```



### Checking initial values

fssg includes a few diagnostic functions for checking whether the init function for a distribution will fail or not based on your data. This simply checks if the initial distribution parameters will allow the distribution function to return valid values for each observed time. 

```{r}
check_inits(
  times = rpois(1000, 100),
  get_fssg_dist('gumbel')
)
```

In the nature of this package, you can also bulk check every distribution on your data! This will perform the simple check_inits function for each distribution packaged within fssg. Please note that this is rather unstable with the flexsurv-native distributions. 

```{r}
bulk_check_inits(rpois(1000,100))
```


## fssg

The titular fssg function provides a one-line way to test multiple distributions for your data. Simply hand it your model formula and/or dataset, and watch it run. By default, the vast majority of the fssg and flexsurv distributions will be run on your data. Progress will be printed, and any models that failed will list the resulting error. The function always prints a summary of all models run, and by default returns a list with two items: the summary table as dataframe, and a list of each successful model.

```{r}
fssg(survival::Surv(time, status) ~ 1, data = survival::cancer) -> output
```


## Evaluating models

fssg comes with a function to extract fit statistics. These statistics are calculated by default for each model within the fssg function, but can be used on individual flexsurv models on its own. 


```{r}
get_fit_stats(Surv(cancer$time, cancer$status), model = output$models$dagum, ibs = F)
```

fssg always comes with a simple function for checking the distributions fit using QQplots. This function creates a simple QQplot of expected quantiles versus observed quantiles. These plots can be helpful for assessing the apropriateness of specific distributions. This function is unstable with flexsurv native distributions.

```{r}
fssg_qqplot(output$models$exp_log)

fssg_qqplot(output$models$gumbel)
```
